A Topic Model of Climate Change Literature

### 2019-07-24 manuscript/coverletter.tex
repository tidied for paper submission

``Big literature'' - dealing with a vast and rapidly expanding body of publications - poses new, and largely neglected challenges for climate change assessments and how we understand them. For example, we estimate a total of 350,000 new publications on climate change to emerge (only) in the Web of Science during the Sixth Assessment Cycle of the Intergovernmental Panel on Climate Change spanning from 2015 to 2022. Big data and machine learning methods need to be applied for keeping an overview of developments in the scientific landscape and to analyse how past IPCC assessments have engaged with the avaailable literature. So far, attempts \cite{Bjurström2011, Hulme2010, Victor2015, Corbera2016, Kowarsch2017} have often been based on qualitiative perceptions, or incomplete data with insufficient methodological sophistication.

In this study we use machine-learning to provide a comprehensive, thematic map of the more than 400,000 articles on climate change published in the Web of Science to date. We perceive two innovations that allow us advance the current state of knowledge on the IPCC and the field it is tasked with assessing. First, our thematic map allows us to situate a complete set of IPCC citations into the context of the wider literature and analyse disciplinary representation adequately.  We find that contrary to previous claims, the social sciences are over-represented in IPCC reports.Second, the use of machine learning methods in creating the map allows us to analyse the content of publications more or less represented in the IPCC. We find that literature on technical solutions (such as negative emissions, buildings or cities) is growing fast and, even compared to other fast-growing literatures, has been under-represented in IPCC reports. We point towards a variety of other applications of such a comprehensive, thematic literature map within global environmetal assessments as well as the broad field of research synthesis. Our findings have direct implications for addressing growing demands for more solution-oriented climate change assessments that are also more firmly rooted in the social sciences.

>>>>

``Big literature'' - the vast and rapidly expanding body of publications - poses new, and largely neglected challenges for climate change assessments and how we understand them. For example, we estimate a total of 350,000 new publications on climate change to emerge (only) in the Web of Science during the Sixth Assessment Cycle of the Intergovernmental Panel on Climate Change spanning from 2015 to 2022. Big data and machine learning methods need to be applied to maintain an overview of developments in the scientific landscape and to analyse how past IPCC assessments have engaged with the available literature. So far, attempts \cite{Bjurström2011, Hulme2010, Victor2015, Kowarsch2017} have often been based on qualitative perceptions, or incomplete data with insufficient methodological sophistication.

In this study we use machine-learning to provide a comprehensive, thematic map of the more than 400,000 articles on climate change published in the Web of Science to date. We perceive two innovations that allow us advance the current state of knowledge on the IPCC and the field it is tasked with assessing. First, our thematic map allows us to situate a complete set of IPCC citations into the context of the wider literature and analyse disciplinary representation adequately.  We find that contrary to previous claims, the social sciences are over-represented in IPCC reports.
Second, the use of machine learning methods in creating the map allows us to analyse the content of publications more or less represented in the IPCC. We find that literature on technical solutions (such as negative emissions, buildings or cities) is growing fast and, even compared to other fast-growing literatures, has been under-represented in IPCC reports. We point towards a variety of other applications of such a comprehensive, thematic literature map within global environmetal assessments as well as the broad field of research synthesis. Our findings have direct implications for addressing growing demands for more solution-oriented climate change assessments that are also more firmly rooted in the social sciences.

### 2019-07-24 manuscript/methods.tex
repository tidied for paper submission

		\begin{table}[htp]
			\begin{center}
				\input{tables/ipcc_matching.tex}
				\caption{The proportion of citations in each report that could be matched with a document from the Web of Science}
				\label{ipcc-matching}
			\end{center}
		\end{table}

@@

	\section*{Supplementary Files}

	\texttt{docs.csv} contains a list of the documents considered in this study, along with basic metadata and their position on the map

@@

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{plots_pub/topic_oecd_entropy.pdf}
			\caption{SI Disciplinary Entropy}
			\label{dis-entropy}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{plots_pub/single_doc_3_536594_1861.pdf}
			\caption{SI Topic make up of a single document}
			\label{doc-topic}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{plots_pub/ipcc_rep_wcs_simplified.pdf}
			\caption{SI Representation by subfield}
			\label{subfield}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{plots_pub/wgs_socsci.pdf}
			\caption{SI Social science \& representation in topics across working groups}
			\label{socsci-wgs}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{plots_pub/topic_rep_ks.pdf}
			\caption{Topic representation over different values of K (number of topics). Topics in the upper or lower 6.66th percentile of either dimension are labelled}
			\label{top-rep-ks}
		\end{center}
	\end{figure}
	\bibliography{Mendeley}

	\bibliographystyle{unsrt}

>>>>

		\begin{table}[htp]
			\begin{center}
				\input{../tables/ipcc_matching.tex}
				\caption{The proportion of citations in each report that could be matched with a document from the Web of Science}
				\label{ipcc-matching}
			\end{center}
		\end{table}

@@

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{../plots_pub/topic_oecd_entropy.pdf}
			\caption{SI Disciplinary Entropy}
			\label{dis-entropy}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{../plots_pub/single_doc_3_536594_1861.pdf}
			\caption{SI Topic make up of a single document}
			\label{doc-topic}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{../plots_pub/ipcc_rep_wcs_simplified.pdf}
			\caption{SI Representation by subfield}
			\label{subfield}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{../plots_pub/wgs_socsci.pdf}
			\caption{SI Social science \& representation in topics across working groups}
			\label{socsci-wgs}
		\end{center}
	\end{figure}

	\begin{figure}
		\begin{center}
			\includegraphics[width=1\linewidth]{../plots_pub/topic_rep_ks.pdf}
			\caption{Topic representation over different values of K (number of topics). Topics in the upper or lower 6.66th percentile of either dimension are labelled}
			\label{top-rep-ks}
		\end{center}
	\end{figure}

	\section*{Data Availability}

	Three datasets from this study are available at \url{https://doi.org/10.6084/m9.figshare.9009665}

	\medskip\noindent
	\texttt{docs.csv} contains a list of the documents considered in this study, along with basic metadata and their position on the map. For copyright reasons, the full metadata from Web of Science can not be published. To reproduce the analysis, it would be necessary to download the abstracts for the papers shown, either using the Web of Science IDs provided, or the query documented in \cite{Grieneisen2011}.

@@

	\medskip\noindent
	\texttt{topic\_comparison.xlsx} shows models with different numbers of topics. It was used to select the topic model used for analysis in this paper.

	\section*{Code Availability}

	The code used to produce this paper is available at https://github.com/mcallaghan/cc-topography

	\bibliography{Mendeley}
	\bibliographystyle{unsrt}
